Comments on \(Q\)-Bosons, and their relations with integrable hierarchies

### 2024-04-12 draft.tex
work on the draft

\title{Comments on \(Q\)-Bosons, deformed fermions and their relations
    with integrable hierarchies}

@@

@@

\begin{abstract}
  In this work, we study aspects of the \(Q\)-boson model, and 
its relations with integrable hierarchies. 
% \vspace{4pt}
 \noindent \textbf{Keywords:} Integrability, Q-bosons, Schur
\end{abstract}

@@

\setcounter{tocdepth}{1}
\tableofcontents

@@

With these expressions, one can finally build the Bethe states and their dual 
\begin{equation}
  \ket{\widehat{\Psi}(y_1, \dots, y_N)} = \prod_{j=1}^N B(y_j) \ket{\bm{0}}\qquad 
  \bra{\widehat{\Psi}(y_1, \dots, y_N)} = \bra{\bm{0}}\prod_{j=1}^N C(y_j) \; .
\end{equation}
When the set \(\{ y_j \ | \ j =1, \dots , N\}\) satisfies the Bethe equations 
\begin{equation}
  y^{N + M}_i = (-1)^{N-1} \prod_{\substack{j = 1 \\ j \neq i}}^N y_j\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we says that the Bethe states are \emph{on-sheel}. In what follows, we will 
only consider \emph{off-shell} states.

\subsubsection{Coordinate expansion}
From the map defined in~\cite{Wheeler:2010vmq}, we can associate to
each state \(\ket{\lambda}\), a Schur function \(s_\lambda(\vec{x})\),
where the dummy coordinates are \(\vec{x} = (x_1, \dots, x_N)\), with
\(N = \sum_{j=0}^M n_j\). Therefore, we can work in the ring of
symmetric functions \(\Lambda_M\), with basis given by Schur functions
\(s_\lambda\) and Young diagrams such that \(\max(\lambda_1) = M\),
and \(\max(\ell(\lambda)) = N\).

That also means that the operator \(B\) should also be realized in the
space \(\Lambda_M\), see~\cite{Tsilevich:2006}. The operators
\(\phi^\dagger_j\) and \(\phi_j\) are, respectively, \(1\)-particle
creation and annihilation at the site \(j\). In terms of partitions,
we can think of them as insertion and removal of a row of length
\(j\). That is, given the state \(\ket{\lambda} =\ket{1^{n_1} 2^{n_2}\cdots}\), 
we have
\begin{equation}
\begin{split}
  \phi^\dagger_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j+1}\cdots} \equiv \ket{\lambda \oplus (j)}\\
  \phi_j \ket{\lambda} & = \ket{1^{n_1} 2^{n_2}\cdots j^{j-1}\cdots}
  \equiv \ket{\lambda \ominus (j)}\; , 
\end{split}
\end{equation}
Once again, we see that the case \(j=0\) is meaningless in the
partition construction. Given two partitions \(\lambda\) and \(\mu\),
we say that \(\lambda\) interlaces \(\mu\) if \(\lambda_i \geq \mu_i \geq \lambda_{i+1}\), 
and we write \(\lambda\succ \mu\). Therefore, it is easy to see that the partition 
\(\lambda\oplus (j)\) interlaces \(\lambda\), and \(\lambda\)
interlaces \(\lambda \ominus (j)\). We write this condition as 
\(\lambda \oplus (j) \succ \lambda \succ \lambda \ominus (j)\).

Now, we can use the explicit expression for the \(L\)-matrix to build
the monodromy matrix, and extract the operator \(B(x)\) as
\begin{subequations}
\begin{equation}
  B(x)_a = T_a(x)_{12} = \sum_{j_1, \dots, j_M = 1,2} L_{a j_M}(x) L_{j_M j_{M-1}}(x)\cdots  L_{j_1 2}(x)\; ,
\end{equation}
and it is easy to see that it has the form  
\begin{equation}
  B(x) \equiv x^{-M/2} \mathbb{B}(x) = x^{-M/2} \sum_{k=0}^M x^k \mathcal{B}_k \; .
\end{equation}
The problem now is to find the action of the operators \(\mathcal{B}_k\) on
states \(\ket{\lambda}\).
\end{subequations}
Let us consider some simple cases.

\textbullet \ M=0: We have just one site in the  
chain, and all particles site at the site \(0\), as we have discussed many times, 
does not contribute to the partition representation.

\textbullet \ M=1: Now we have 1 row Young diagrams \(\lambda = (1^{n_1})\), where each row
corresponds to a boson. It is easy to show that 
\begin{equation}
 \mathbb{B}_{M=1}(x) =  x \phi_1^\dagger + \phi_0^\dagger \; ,
\end{equation}
therefore, \(\mathcal{B}_0 = \phi_0^\dagger\) and  \(\mathcal{B}_1 = \phi_1^\dagger\). 
From this expression, we have 
\begin{equation}
  \mathcal{B}_0\ket{(1^{n_1})} = \ket{(1^{n_1})} \qquad  
  \mathcal{B}_1\ket{(1^{n_1})} = \ket{(1^{n_1+1})} \; .
\end{equation}
All in all, we conclude that 
\begin{equation}
 \mathbb{B}_{M=1}(x)\ket{\lambda}  =  \ket{\lambda} + x\ket{\lambda \oplus (1)} \;.
\end{equation}

\textbullet \ M=2: Now we have 2 row Young diagrams \(\lambda = (1^{n_1}2^{n_2})\).
Now, one can show that
\begin{equation}
  \mathbb{B}_{M=2}(x) = x^2 \phi_2^\dagger + x \left( \phi_1^\dagger
  + \phi_2^\dagger \phi_1 \phi_0^\dagger \right) + \phi_0^\dagger \; ,
\end{equation}
therefore,
\begin{equation}
\mathcal{B}_0 = \phi_0^\dagger \qquad 
\mathcal{B}_1 = \phi_1^\dagger + \phi_2^\dagger \phi_1 \phi_0^\dagger \qquad 
\mathcal{B}_2 = \phi_2^\dagger \; .
\end{equation}

Consequently. we find
\begin{subequations}
\begin{equation}
\begin{split}
  \mathcal{B}_0\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1} 2^{n_2})} \qquad  
  \mathcal{B}_1\ket{(1^{n_1}2^{n_2})} = \ket{(1^{n_1+1} 2^{n_2})} + \ket{(1^{n_1-1} 2^{n_2 + 1})} \\
  \mathcal{B}_2\ket{(1^{n_1}2^{n_2})} & = \ket{(1^{n_1}2^{n_2+1})}\; .
\end{split}
\end{equation}
\end{subequations}
Observe that the action of the operator \(\mathbb{B}_k(x)\) on \(\ket{\lambda}\)
generates diagrams \(\ket{\mu}\) such that \(k = |\mu| - |\lambda|\). 
\begin{equation}
  \mathbb{B}_{M=2}(x) \ket{\lambda}  =
  \sum_{\lambda \prec \mu \subseteq [2,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
\end{equation}

One can keep doing this game, and we reach to the general conclusion that 
\begin{equation}
  \mathbb{B}(x) \ket{\lambda}  =
  \sum_{\lambda \prec \mu \subseteq [M,\ell(\lambda) + 1]} x^{|\mu|-|\lambda|} \ket{\mu}\; .
\end{equation}

Therefore, we conclude that the action of the operators
\(\mathcal{B}_k\) on \(\ket{\lambda}\) generates diagrams
\(\ket{\mu}\) that are obtained from \(\lambda\) by adding \(k\) boxes,
not in different columns. From the viewpoint of symmetric functions, this rule
is exactly the Pieri's rule~\cite{Macdonald:1998}, that states that
we have that
\begin{equation}
  s_{(k)} s_\lambda = \sum_{\mu \succ \lambda} s_\mu\;.
\end{equation}

Therefore, we have the map from the spin chain Hilbert space to the
ring of symmetric functions given by \(\imath : \mathcal{B}_k \mapsto
s_{(k)}\) where \(s_{(k)}\) is the Schur function with 1 row of length
\(k\).

@@

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Toda tau functions and the phase model}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

Bogoliubov~\cite{Bogoliubov2005}, in the notation~\cite{Wheeler:2010vmq},
has shown that the the scalar product of two vectors in the
\(N\)-particles sector in a chain of length \(M + 1\) is
\begin{equation}
\begin{split}
\label{eq:scalar}
  S(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\) is an \(N\times N\) matrix with components
\begin{equation}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

This scalar product is related to the counting of plane
partitions, with
\begin{equation}
 S(N, M| \bm{x}, \bm{y}) 
 = \sum_{\lambda \subseteq (N, N, M)} s_{\lambda}(\bm{x}) s_{\lambda} (\bm{y})\; .
\end{equation}
In the limit \(M, N \to \infty\), one can use the Cauchy identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{x}) s_{\lambda}(\bm{y})
  = \prod_{i,j>0} \frac{1}{1- x_iy_j}\; ,
\end{equation}
and specialize this expression to \( x_i = y_i = q^{i-1/2} \) to find
the MacMahon function
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{q}_\star) s_{\lambda}(\bm{q}_\star) = \prod_{i,j>0} \frac{1}{1- q^{i +j -1}}\; ,
\end{equation}
where \(\bm{q}_{\star} = (q^{1/2}, q^{3/2}, \dots)\).

\subsection{KP and Toda tau functions}

Wheeler has shown in~\cite{Wheeler:2010vmq} that the the scalar
product~(\ref{eq:scalar}) is a KP tau function with respect to the
Miwa coordinates \(t_m = \frac{1}{m}\sum_j x_j^m\) (\(t'_m =
\frac{1}{m}\sum_j y_j^m\)), where the Schur functions
\(s_{\lambda}(\bm{y}) = c_\lambda\) ( \(s_{\lambda}(\bm{x}) =
c'_\lambda\)) become Plücker coordinates, and vice-versa.
The proof for this statement is relies on the map \(\mathcal{M}_\psi\)
between the space \(\mathcal{V}\) os Bethe states and Fock
space \(\mathcal{F}_\psi\) of free fermions.

In fact, there is an easier way to see this fact. First one may see that 
\begin{equation}
 S(N,M,\bm{y}|\{\bm{t}\}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) 
\end{equation}
is a tau function of the KP hierarchy, since
\(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\) are Plücker
coordinates~\cite{Miwa2000, Foda:2009zz}. Therefore, the truncation
\begin{equation}
 S(N,M,\bm{y}|\{\bm{t}\}) 
 = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
\end{equation}
is also a tau function. According to~\cite{Alexandrov:2012tr,
  Kharchev:1991gd, Zabrodin:2010ii}, the truncation of the tau
function corresponds to the inclusion of a projection operator
\(\mathrm{P}^+\) in the expectation value of the tau function.
% See section 3.3.2 of~\cite{Alexandrov:2012tr},

\subsubsection{Scalar product}
We also argue that this is also a tau function of the 2D toda hierarchy. 
We can write the Cauchy's identity as
\begin{subequations}
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_m \right) \; .
\end{equation}
where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.
It is a trivial tau function of the 2D Toda hierarchy, since one can write it as
\begin{equation}
\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}
\end{equation}
and from the free-fermions representation of the tau functions,
see~\cite{Alexandrov:2012tr}, we have that it is a tau function of the
Toda hierarchy with trivial element \(1 \in GL(\infty)\).
\end{subequations}

Therefore, the truncation 
\begin{equation}
 S(N,M|\{\bm{t}\},\{\bm{t}'\}) = 
  \sum_{\lambda \subseteq [N,N,M]} s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
\end{equation}
is a also a restricted tau function. These proofs follow from simple
properties of tau functions, and it should be contrated with the rich
formalism in~\cite{Wheeler:2010vmq}. Moreover, from~\cite{Zabrodin:2010ii}
we have that the this tau function can be written as the matrix integral
\begin{equation}
  S(N,M|\{\bm{t}\}, \{\bm{t}'\}) =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi_+(\bm{t}_+, z_\ell) - \xi_-(\bm{t}_-, z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})
\end{equation}

@@

\subsubsection{Correlation functions}
Bogoliubov has also shown that the correlation functions
\begin{subequations}
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \bra{0} \prod_{j=1}^N C(x) \prod_{j=1}^{N-1} B(y) \phi_m^\dagger \ket{0}
\end{equation}
can be written as
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \frac{(-1)^{N-1}}{(\sqrt{y_N}^{N-1}}
  \left( \prod_{t<N} \frac{x_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where the \(N\times N\) matrix \(Q\) has components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\). 
Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
we cannot write the above expression in terms of a Toda hierarchy tau
function.

@@

%\addcontentsline{toc}{section}{References}
\printbibliography

>>>>

\title[\(Q\)-Bosons and relations with integrable hierarchies]{Comments on \(Q\)-Bosons,
and their relations with integrable hierarchies}

@@

\begin{abstract}
In this work, we study aspects of the \(Q\)-boson model, and 
its relations with integrable hierarchies. 
\vspace{4pt}
\noindent \textbf{Keywords:} Integrability, Q-bosons, Schur

@@

\end{abstract}

@@

\setcounter{tocdepth}{3}
\tableofcontents

@@

With these expressions, one can finally build the Bethe states
\begin{equation}
  \ket{\Psi(y_1, \dots, y_N)} = \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}}
\end{equation}
where \(\mathbb{B}(x) = x^{M/2} B(x)\). When the set \(\{ y_j \ | \ j
=1, \dots , N\}\) satisfies the Bethe equations
\begin{equation}
  y^{N + M}_i = (-1)^{N-1} \prod_{\substack{j = 1 \\ j \neq i}}^N y_j\; , \qquad i = 1, \dots, N\; , 
\end{equation}
we says that the Bethe states are \emph{on-sheel}. In what follows, we will 
only consider \emph{off-shell} states.

@@

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*
\section{Tau functions in the phase model}
%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

This section discusses presence of integrable hierarchies in the phase
model.  We argue that it is also a Toda hierachy tau function, and we
discuss some consequences of this fact, in particular, its connection
with a matrix model. Moreover, we also point the presence of
correlation functions in this model that also satisfy the KP hierarchy
equations. We finally argue how these results also imply that these
scalar products are also KP hierarchy tau functions, what is expected from 
the results of Wheeler~\cite{Wheeler:2010vmq}.

Bogoliubov~\cite{Bogoliubov2005}, has shown that the scalar product of
two vectors in the \(N\)-particles sector in a chain of length \(M+1\)
is
\begin{equation}
\begin{split}
\label{eq:scalar}
  S(N, M|\bm{x}, \bm{y}) & =
  \bra{\bm{0}} \prod_{j=1}^N \mathbb{C}(x_j) \prod_{j=1}^N \mathbb{B}(y_j) \ket{\bm{0}} \\ 
 & = \frac{ \det H(\bm{x},\bm{y})}{ \prod_{i<j}(x_i - x_j)(y_i - y_j)} \; ,
\end{split}
\end{equation}
where \(H\) is an \(N\times N\) matrix with components
\begin{equation}
  H_{ij} = H(x_i, y_j) 
  =\frac{1 - (x_i y_j)^{ M + N}}{1 - x_i y_j }\; .
\end{equation}

\subsection{Toda tau functions}
In this section, we argue that the scalar product defined above is a tau function 
of the Toda hierarchy. Let us first write the function \(H(z,w)\) as the geometric sum 
\begin{equation}
  H(z,w) = \frac{1 - (zw)^{M+N}}{1 - zw} = \sum_{k=1}^{M+N} (zw)^{k-1} \; , 
\end{equation}
in such a way that 
we can write
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) = \det_{i,j} \left( \sum_{k=1}^{M+N} x_i^{k-1} y_j^{k-1}\right)
\end{equation}
and we can thing of this expression as the product of a \(N\times (N+M)\) matrix \(\mathcal{X}\)
and another \((M + N)\times N\) matrix \(\mathcal{Y}\), given by
\begin{equation}
  \mathcal{X} = 
  \begin{pmatrix}
  x_1^0 & x_1^1 & \dots & x_1^{M+N-1} \\  
  x_2^0 & x_2^1 & \dots & x_2^{M+N-1} \\  
  \vdots \\
  x_N^0 & x_N^1 & \dots & x_N^{M+N-1} 
  \end{pmatrix}\quad \textrm{and} \quad 
  \mathcal{Y} = 
  \begin{pmatrix}
  y_1^0 & y_2^0 & \dots & y_N^0 \\  
  y_1^1 & y_2^1 & \dots & y_N^1 \\  
  \vdots & \vdots & & \vdots \\
  y_1^{N+M-1} & y_2^{M+N-1} & \dots & y_N^{M+N-1} 
  \end{pmatrix}
\end{equation}
therefore 
\begin{equation}
  \det_{i,j} \left(H(x_i, y_j)\right) =  \det_{i,j} \left( \mathcal{X}\mathcal{Y}\right)
= \sum_{0 \leq \ell_{N} \leq \dots \leq \ell_1\leq N+M }  \det_{ik}(x_i^{\ell_k})  \det_{ik}(y_j^{\ell_k})\; .
\end{equation}
If we now define \(\ell_j = \lambda_k - k + N\), and
using~(\ref{eq:scalar}), we have
\begin{equation}
S(N, M|\bm{x}, \bm{y}) = \sum_{\lambda\subseteq [N,M]}
\frac{\det_{ik}(x_i^{\lambda_k - k + N})}{\Delta(\bm{x})} \frac{ \det_{ik}(y_j^{\lambda_k - k + N})}{\Delta(\bm{y})}
= \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{x}) s_\lambda(\bm{y}) \; .
\end{equation}
where \(\Delta(\bm{x})\) and \(\Delta(\bm{x})\) are Vandermonde
determinants. This formula agrees with the Schur expansion defined
in~\cite{Bogoliubov2005}.

Let us write the Miwa coordinates 
\begin{equation}
  t_p = \frac{1}{p}\sum_{j=1}^N x_j^p\qquad 
  t'_p = \frac{1}{p}\sum_{j=1}^N y_j^p\; ,
\end{equation}
therefore, we have 
\begin{equation}
  S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
In fact, using the free fermions representation~\cite{Alexandrov:2012tr},
it can be written as
\begin{equation}
\sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element \(1 \in GL(\infty)\).
In fact, it is nothing but the Cauchy's identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_m \right) \; .
\end{equation}
where \(\bm{t}\) and \(\bm{t}'\) are two independent sets of Miwa coordinates.

Putting all these facts together, the truncation for finite \(M\) and
\(N\) are also tau functions of the Toda hierarchy. More specifically,
according to~\cite{Alexandrov:2012tr, Kharchev:1991gd,
  Zabrodin:2010ii}, the truncation of the tau function corresponds to
the inclusion of a projection operator \(\mathrm{P}^+\) in the
expectation value of the tau function written in the fermionic
representation.

\subsubsection{KP tau function}
Finally, one may also observe that if we fix one set of coordinates,
say \(\bm{y}\), then we write
\begin{equation}
 S(N,M|\{\bm{t}\}) 
 = \sum_{\lambda} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
 \equiv \sum_{\lambda} c_\lambda(\bm{y})  s_\lambda(\bm{t}) \; ,
\end{equation}
with coefficients \(c_\lambda(\bm{y}) = \det (h_{\lambda_i-i +j}(\bm{y}))\).
But now, it is a trivial to notice that these are Plücker
coordinates in the Jacobi-Trudi form, see~\cite{Miwa2000, Alexandrov:2012tr}.

Therefore, the expression
\begin{equation}
 S(N,M|\{\bm{t}\}) = \sum_{\lambda \subseteq [N, \infty)} s_\lambda(\bm{y})  s_\lambda(\bm{t}) 
\end{equation}
is also a KP tau function: A fact that we already know
from~\cite{Wheeler:2010vmq}, where the author has proved this
statement using the free fermions formalism.

\subsection{Matrix models}

Moreover, from~\cite{Zabrodin:2010ii} we have that the this tau
function can be written as the matrix integral
\begin{equation}
  S(N,M|\{\bm{t}\}, \{\bm{t}'\}) =
  \frac{1}{N!} \prod_{\ell=1}^N \oint_{\Gamma_\ell} \frac{dz_\ell}{2 \pi i z_\ell}
  e^{\xi_+(\bm{t}_+, z_\ell) - \xi_-(\bm{t}_-, z_{\ell}^{-1})} \Delta(z)\Delta(z^{-1})
\end{equation}

@@

In fact, we can also write
\begin{equation}
  H(z,w) = \sum_{k=0}^{M+N} h_k(zw)^{k-1} \; , 
\end{equation}
with \(h_k=1\) if \(k\in [0, M+N-1]\) and \(0\) otherwise. In this
case, one can define a diagonal \((N+M)\times (N+M)\) matrix
\(\mathcal{H} = \textrm{diag}(h_0, \dots, h_{M+N-1})\). Therefore, if
we repeat the arguments above, we find
\begin{equation}
S(N, M|\bm{t}, \bm{t}') = \sum_{\lambda} h_\lambda s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
where \(h_\lambda = 1\) if \(\lambda \subset [N,M]\) and it vanishes otherwise.

\subsection{Correlation functions}
Bogoliubov has also shown that the correlation functions
\begin{subequations}
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \bra{0} \prod_{j=1}^N C(x) \prod_{j=1}^{N-1} B(y) \phi_m^\dagger \ket{0}
\end{equation}
can be written as
\begin{equation}
  P(m|\bm{x}, \bm{y}) = \frac{(-1)^{N-1}}{(\sqrt{y_N}^{N-1}}
  \left( \prod_{t<N} \frac{x_N - y_t}{y_t} \right) \frac{\det Q}{\det H} S(N,M|\bm{x}, \bm{y})
\end{equation}
\end{subequations}
where the \(N\times N\) matrix \(Q\) has components 
\begin{equation}
 Q_{jN} = x_j^{(M + N - 1- 2m)/2} \quad  \textrm{and} \quad 
 Q_{jk} = H_{jk} \; , 
\end{equation}
where \(H_{jk} = H(x_j, y_k)\) are the components of the matrix \(H\). 
Since the components \(Q_{jN}\) are independent of the coordinates \(y\), 
we cannot write the above expression in terms of a Toda hierarchy tau
function.

@@

%\addcontentsline{toc}{section}{References}

\printbibliography
